A Structural and Functional Bioinformatics Study of QTY-designed Retinylidene Proteins

### 2025-06-09 Paper/Manuscript-revised/manuscript.tex
Revise manuscript

Retinylidene proteins are retinal-binding light-sensitive proteins found in organisms ranging from microbes to human. Microbial opsins have been utilized in optogenetics, while animal opsins are essential for vision and light-dependent metabolic functions. However, retinylidene proteins have hydrophobic transmembrane (TM) domains, which makes them challenging to study. In this structural and functional bioinformatics study, I used the QTY (Glutamine, Threonine, Tyrosine) code to design water-soluble QTY analogs of retinylidene proteins, including 9 human and 3 microbial opsins. I provide superpositions of the AlphaFold3-predicted hydrophobic native proteins and their water-soluble QTY analogs, and experimentally-determined structures when available. I also provide a comparison of surface hydrophobicity of the variants. Despite significant changes to the protein sequence (35.53\% to 50.24\% in the TM domain), protein characteristics and structures are well-preserved. Furthermore, I run molecular dynamics (MD) simulations of native and QTY-designed OPN2 (rhodopsin) and analyze their response to the isomerization of 11-cis-retinal to all-trans-retinal. The results show that the QTY analog has similar functional behavior to the native protein. The findings of this study indicate that the QTY code can be used as a robust tool to design water-solubility retinylidene proteins. These have potential applications in protein studies, therapeutic treatments, and bioengineering.

@@

Retinylidene proteins are photochemically reactive proteins that are bound to or can bind to retinal (vitamin A aldehyde) as their chromophore \citep{Spudich_2000}. I will use the term ``opsin'' interchangeably with ``retinylidene protein'', despite the fact that ``opsin'' sometimes refers specifically to the chromophore-free apoprotein. Retinylidene proteins are divided into two groups, animal opsins and microbial opsins, which are evolutionarily distinct but share common characteristics such as having 7 transmembrane (TM) domains and a retinal-binding lysine residue \citep{Yee_2013, Spudich_2000}. In this study, I will consider 9 animal opsins and 3 microbial opsins.

\textbf{Animal opsins} are a type of class A GPCR (G-protein coupled receptor), which are characterized by 7 transmembrane domains, the NPxxY motif, and activation of G-proteins through the outward movement of TM6. \citep{Zhou_2019}. Almost all animal opsins include a lysine residue that forms a Schiff base link with the retinal \citep{Guhmann_2022}. When retinal absorbs a photon, it isomerizes, usually changing from 11-cis to all-trans. The subsequent conformational changes of the protein are well studied using bovine rhodopsin, which changes from dark state to BATHO state, then LUMI, META I, and META II \citep{Okada_2001}. Proton transfer plays an important role in this process \citep{Mahalingam_2008}. As the protein conformation changes, TM6 characteristically moves outward, activating the G-protein.

In this study, I selected 9 opsins expressed in the human nervous system: OPN1MW (UniProt ID: \href{https://www.uniprot.org/uniprotkb/P04001/entry}{P04001}), OPN1LW (UniProt ID: \href{https://www.uniprot.org/uniprotkb/P04000/entry}{P04001}), OPN1SW (UniProt ID: \href{https://www.uniprot.org/uniprotkb/P03999/entry}{P03999}), OPN2 (UniProt ID: \href{https://www.uniprot.org/uniprotkb/P08100/entry}{P04001}), OPN3 (UniProt ID: \href{https://www.uniprot.org/uniprotkb/Q9H1Y3/entry}{Q9H1Y3}), OPN4 (UniProt ID: \href{https://www.uniprot.org/uniprotkb/Q9UHM6/entry}{Q9UHM6}), OPN5 (UniProt ID: \href{https://www.uniprot.org/uniprotkb/Q6U736/entry}{Q6U736}), RGR (UniProt ID: \href{https://www.uniprot.org/uniprotkb/P47804/entry}{P47804}) and RRH (UniProt ID: \href{https://www.uniprot.org/uniprotkb/O14718/entry}{O14718}). They belong to several evolutionarily distinct families of animal opsins \citep{Terakita_2005}.

@@

\textbf{OPN2} (Opsin 2), also known as rhodopsin, is expressed in retinal rod photoreceptors and is responsible for vision at low light intensity \citep{Hubbard_1958}. Certain variants lead to autosomal recessive or autosomal dominant retinitis pigmentosa and congenital stationary night blindness \citep{Fanelli_2021}. OPN2 is a representative animal opsin, one of the first studied. In fact, bovine OPN2 is the first opsin to be sequenced \citep{Nathans_1984}, as well as the first GPCR whose crystal structure was resolved experimentally \citep{Palczeski_2000}. Many studies on the functional mechanisms of animal opsins also focus on OPN2. Consequently, I choose OPN2 to conduct a functional analysis in this study, in order to further explore the effectiveness of the QTY code in redesigning retinylidene proteins.

@@

In this study, I selected 3 microbial opsins: BACR (UniProt ID: \href{https://www.uniprot.org/uniprotkb/P02945/entry}{P02945}), BACH (UniProt ID: \href{https://www.uniprot.org/uniprotkb/B0R2U4/entry}{B0R2U4}), ChR2 (UniProt ID: \href{https://www.uniprot.org/uniprotkb/Q8RUT8/entry}{Q8RUT8}).

@@

Instead of taking a computational approach, I applied the QTY (Glutamine, Threonine, Tyrosine) code to systematically engineer water-soluble analogs with reduced hydrophobicity in membrane proteins. There are structural similarities between hydrophobic and polar amino acids: leucine (L) and glutamine (Q); isoleucine (I) / valine (V) and threonine (T); and phenylalanine (F) and tyrosine (Y), as can be observed on high-resolution electron density maps \citep{Zhang_2018, Zhang_2022, Tegler_2020}. This justifies the replacement of hydrophobic amino acids with polar ones. Zhang et al. initially applied the QTY code to design several detergent-free chemokines receptors, all of which retained structural thermal stability and native ligand-binding and enzymatic activities despite substantial changes to the transmembrane domain \citep{Zhang_2018, Tegler_2020}. Zhang's team have also applied the QTY code to design water-soluble GPCRs, including chemokine receptors \citep{Zhang_2018, Qing_2019, Tegler_2020, Skuhersky_2021}, cytokine receptors \citep{Hao_2020}, and olfactory receptors \citep{Skuhersky_2021, Johnsson_2025}. In addition to GPCR, they have also used the QTY code to design water-soluble glucose transporters \citep{Smorodina_2022}, ABC transporters \citep{Pan_2024}, monoamine neurotransmitter transporters \citep{Karagol_Karagol_Alper_2024}, glutamate transporters \citep{Karagol_Karagol_Smorodina_2024, Karagol_2024}, mitochondrial megacomplex \citep{Chen_2025}, antibodies \citep{Li_Wang_2024}, potassium channels \citep{Smorodina_2024}, receptor kinases \citep{Li_Tang_2024}, transmembrane enzymes \citep{Chen_Pan_2025}, as well as bacterial enzymes with beta barrel structures \citep{Sajeev-Sheeja_2023, Sajeev-Sheeja_2024}. The QTY-designed water-soluble CXCR4 chemokine receptor has been successfully used in biomimetic sensors \citep{Qing_2023}. They have also applied the rQTY (reverse QTY) code to design self-assembled nanoparticles \citep{Meng_2023}.

AlphaFold2 was released by Google DeepMind in July 2021 \citep{Jumper_2021}. It has greatly facilitated the study of QTY-designed protein variants. Subsequently, AlphaFold3 was released in May 2024, featuring improved architecture and improved efficiency \citep{Abramson_2024}. Furthermore, AlphaFold3 enabled the accurate prediction of complexes of multiple proteins, as well as complexes with modified residues, nucleic acids, ions, certain ligands. QTY studies have made use of these new features \citep{Chen_2025, Johnsson_2025}.

@@

I predicted the structures of native proteins and their QTY variants using the AlphaFold3 website (\url{https://alphafoldserver.com}) \citep{Abramson_2024}. For microbial opsins, the dimers and trimers had identical subunits, i.e., they were homodimers and homotrimers. In AlphaFold3, this was achieved by altering the number of protein copies.

@@

PDB files for native protein structures determined experimentally by X-ray diffraction or electron microscopy were taken from RCSB PDB: OPN2 (PDB ID: \href{https://www.rcsb.org/structure/5W0P}{5W0P}), BACR (PDB ID: \href{https://www.rcsb.org/structure/7XJC}{7XJC}), BACH (PDB ID: \href{https://www.rcsb.org/structure/2JAF}{2JAF}), ChR2 (PDB ID: \href{https://www.rcsb.org/structure/8ZAN}{8ZAN}). The AlphaFold3-predicted native and QTY variants were taken directly from the most probable predicted structure. Superpositions were performed via the command ``super'' in PyMOL (\url{https://pymol.org}). I remove unstructured loops at the N and C terminals for the sake of clarity.

@@

Regarding native OPN2, the membrane protein system was constructed using the web-based membrane builder CHARMM-GUI and was downloaded in GROMACS format \citep{Jo_2008, Wu_2014, Lee_2016}. The protein was centered in a rectangular box. The generated membrane models consisted of 40\% POPC, 40\% POPE, 10\% POPS, and 10\% cholesterol, which simulate a rod photoreceptor disk membrane \citep{Albert_2005}. The system was solvated in TIP3P water with 150 mM NaCl. 11-cis-retinal was manually added to the protein files. LINCS constraints were used for constraints, and Verlet integrator was used. Electrostatics was handled with Particle Mesh Ewald (PME), with both Coulomb and van der Waals interaction cutoffs set at 1.2 nm. The energy of the system was minimized using the steepest descent until the maximum forces converged below 1000 kJ/mol/nm. The standard six-step CHARMM-GUI NP$\gamma$T equilibration protocol \citep{Jo_2008} was used, with two 125-ps NVT equilibration simulations, and one 125-ps and three 500-ps NP$\gamma$T equilibration simulations. Temperature and pressure were maintained at 303.15 K and 1.0 bar, respectively, using the V-rescale thermostat and C-rescale barostat with surface-tension coupling. Following equilibration, a 10-ns production MD simulation was run. The retinal molecule was then manually switched to the all-trans state by rotating C13 to C15 and bonded hydrogen and methyl groups by 180 degrees around the C11-C12 bond. This mimics the effect of the absorption of a photon by 11-cis-retinal. Energy minimization and equilibration were repeated exactly as before, and then a 120-ns production MD simulation was run.

Regarding QTY-designed OPN2, I found that the helix 9 of the QTY analog was flung up and adhered to the originally transmembrane domain, which disrupted the normal conformation. Therefore, this helix was truncated. The system was constructed directly using GROMACS. The protein was centered in a rectangular box with dimensions equal to those of native OPN2. The equilibration involved a sequence of simulations: one 250-ps NVT equilibration, followed by one 125-ps and one 1500-ps NPT equilibrations, totaling an equilibration duration that matches that of the natural OPN2. All other configurations and parameters are identical to those of the original OPN2.

@@

The transmembrane segments of membrane proteins are hydrophobic in nature. This leads to a challenge to study their structure and function. Traditionally, detergent must be used to purify them. Zhang et al. considered a different approach, i.e., via systematic soluble protein design. There exists structural similarities, as can be observed on high-resolution electron density maps, between certain hydrophobic and hydrophilic amino acids: leucine (L) and glutamine (Q), isoleucine (I) or valine (V) and threonine (T), and phenylalanine (F) with tyrosine (Y). This fact enables systematic replacement of L with Q, I/V with T, and F with Y in all transmembrane segments of the membrane protein, which Zhang et al. named the QTY code. QTY analogs are less hydrophobic than native membrane proteins. Although their amino acid sequences have been significantly changed, they still exhibit relatively preserved structure, isoelectric points (pI), and molecular weights (MW) compared to the native membrane proteins (Table~\ref{tb:characteristics}).

@@

The protein sequences of 9 human opsins and 3 microbial opsins are aligned with their QTY analogs (Figure~\ref{fig:sequences}). The QTY substitution resulted in an overall variation of the sequence of 15.48\% to 30.04\% and a variation of transmembrane domain of 35.53\% to 50.24\%. Despite changes in amino acid composition and sequence, the pI only experiences a slight change, between 0.00 and 0.16. This is because the amino acids Q, T, and Y have neutral charges and do not introduce additional charges to the system.  The MW increases slightly by a value between 0.04kDa and 0.60kDa. This is because the hydrophilic amino acids usually have nitrogen or oxygen atoms in place of carbon atoms in hydrophobic acids and thus may have higher mass: L (131.17Da) \textit{vs} Q (146.14Da); I (131.17Da) and V (117.15Da) \textit{vs} T (119.12Da); F (165.19Da) \textit{vs} Y (181.19Da).

@@

Although the chromophore, retinal, is bound to retinylidene proteins in an internal pocket, it is only incorporated into the protein after protein folding is complete. In other words, protein folding does not depend on the chromophore. Thus, it is justified to predict the ligand-free forms of the proteins. I used AlphaFold3 to predict the structure of the 9 native human opsins and their water-soluble QTY analogs. The structures superpose very well (Figure~\ref{fig:humansup} \textbf{a)} to \textbf{i)}). For OPN2, which has an available X-ray diffraction structure, I also superposed the experimentally determined structure with the AlphaFold3-predicted structures. Overall, the root mean square distances (RMSD) are small, from 0.307{\AA} to 0.611\AA, with only one exception of OPN2$^{\textrm{QTY}}$ \textit{vs} OPN2$^{\textrm{EXP}}$ (RMSD = 0.999\AA). This shows that the water-soluble QTY analogs are quite similar to native proteins in terms of structure.

All 9 of these human opsins belong to the same large family, i.e. animal opsins. However, they belong to different subfamilies and have structural differences from each other. In order to investigate the extent to which the water-soluble QTY analogs preserve these differences, I also conducted pairwise superpositions within all 9 native human opsins and within all 9 QTY analogs of human opsins (Figure~\ref{fig:humansup} \textbf{j)} and \textbf{k)}). As observed in the tables, there is a large degree of similarity between the RMSD between a pair of native proteins and that between the corresponding pair of QTY analogs. Therefore, I conclude that the structural differences are relatively well-preserved before and after the QTY substitution. However, whether the functional differences are preserved remains a problem to study.

@@

I used AlphaFold3 to predict the structure of monomers of the 3 native microbial opsins, their water-soluble QTY analogs. Since microbial opsins often have to form oligomers to be functional, I also predicted the structure of native and QTY variants of BACR trimers, BACH trimers, and ChR2 dimers. These predicted structures are also superposed with experimentally determined structures (X-ray diffraction or electron microscopy). The structures superpose very well (Figure~\ref{fig:microbialsup}). Overall, the root mean square differences (RMSD) are small, with the highest value being 0.685\AA. This shows that the water-soluble QTY analogs are likely capable of forming oligomers with structures similar to those of native proteins.

@@

The hydrophobic portions of the protein surface are represented in yellow (Figure~\ref{fig:hydrophobicity}). In the native proteins, there are large hydrophobic patches in the transmembrane domains, which are embedded within the lipid bilayer. The nonpolar and hydrophobic amino acids including leucine (L), isoleucine (I), valine (V), phenylalanine (F), alanine (A), methionine (M), and tryptophan (W) interact with the hydrophobic hydrocarbon chains and expel water.

I investigate the surface hydrophobicity of water-soluble QTY-analogs in comparison with the native proteins. After applying the QTY code to respectively replace the hydrophobic amino acids L, I/V, and F with hydrophilic amino acids glutamine (Q), threonine (T), and tyrosine (Y), the hydrophobic surface areas are significantly reduced. More importantly, since the electron density maps of the corresponding amino acids are similar, the alpha helices in QTY analogs have a similar contour to those in the native proteins. The QTY analogs have successfully retained its structural integrity and stability.

@@

AlphaFold3 is easily accessible online (\url{https://alphafoldserver.com}), allowing users to make 30 predictions a day. The structures of the QTY analogs were predicted using the AlphaFold3 server, which was run free of charge and the results were produced within a few minutes.

DeepMind also collaborates with the EBI to make over 214 million predicted protein structures available through the AlphaFold Protein Structure Database (\url{https://alphafold.ebi.ac.uk}). This number is continuously expanding, with the quality of predictions further improving with the advent of AlphaFold 3.

However, AlphaFold3 still has limitations that have been encountered in this study. Although AlphaFold3 allows the structural prediction of complexes that include small-molecule ligands, it only supports a small range of ligands, which does not include retinal. I were therefore unable to predict the structure of chromophore-bound states of retinylidene proteins with AlphaFold3. Consequently, I chose another approach, namely molecular dynamics simulation, in order to investigate the behavior of QTY opsin analogs when they are bound to the chromophore.

@@

QTY substitutions do not introduce changes to the essential NPxxY motif or the retinal-binding lysine of human retinylidene proteins. Therefore, I propose that QTY analogs of human opsins may have conserved functions and carry out a molecular dynamics (MD) simulation. I record and analyze the behavior of native OPN2 and its QTY analog before and after the 11-cis to all-trans isomerization.

The MD simulation provides evidence of conformational change in both native and QTY analogs (Figure~\ref{fig:conformation}). The native protein experiences a slower change from its original conformation, while the QTY analog shows more fluctuations and sudden changes. Under further inspection, this phenomenon may be due to the entrance of water molecules into the QTY analog. In addition, retinal molecule has a greater degree of change in conformation in the native protein than in the QTY analog. Nevertheless, the retinal in the QTY analog is still able to cause changes to surrounding residues, sometimes doing so through contact with water molecules.

>>>>

Retinylidene proteins are retinal-binding light-sensitive proteins found in organisms ranging from microbes to human. Microbial opsins have been utilized in optogenetics, while animal opsins are essential for vision and light-dependent metabolic functions. However, retinylidene proteins have hydrophobic transmembrane (TM) domains, which makes them challenging to study. In this structural and functional bioinformatics study, I use the QTY (Glutamine, Threonine, Tyrosine) code to design water-soluble QTY analogs of retinylidene proteins, including 9 human and 3 microbial opsins. I provide superpositions of the AlphaFold3-predicted hydrophobic native proteins and their water-soluble QTY analogs, and experimentally-determined structures when available. I also provide a comparison of surface hydrophobicity of the variants. Despite significant changes to the protein sequence (35.53\% to 50.24\% in the TM domain), protein characteristics and structures are well-preserved. Furthermore, I run molecular dynamics (MD) simulations of native and QTY-designed OPN2 (rhodopsin) and analyze their response to the isomerization of 11-cis-retinal to all-trans-retinal. The results show that the QTY analog has similar functional behavior to the native protein. The findings of this study indicate that the QTY code can be used as a robust tool to design water-solubility retinylidene proteins. These have potential applications in protein studies, therapeutic treatments, and bioengineering.

@@

Retinylidene proteins are photochemically reactive proteins that are bound to or can bind to retinal (vitamin A aldehyde) as their chromophore \citep{Spudich_2000}. For convenience, I will use the term ``opsin'' interchangeably with ``retinylidene protein'', despite the fact that ``opsin'' sometimes refers specifically to the chromophore-free apoprotein. Retinylidene proteins are divided into two groups, animal opsins and microbial opsins, which are evolutionarily distinct but share common characteristics such as having 7 transmembrane (TM) domains and a retinal-binding lysine residue \citep{Yee_2013, Spudich_2000}. In this study, I will consider 9 animal opsins and 3 microbial opsins.

\textbf{Animal opsins} are a type of class A GPCR (G-protein coupled receptor), which are characterized by 7 transmembrane domains, the NPxxY motif, and activation of G-proteins through the outward movement of TM6. \citep{Zhou_2019}. Almost all animal opsins include a lysine residue that forms a Schiff base link with the retinal \citep{Guhmann_2022}. When retinal absorbs a photon, it isomerizes, usually changing from 11-cis to all-trans. The subsequent conformational changes of the protein are well studied using bovine rhodopsin, which changes from dark state to BATHO state, then LUMI, META I, and META II \citep{Okada_2001}. Proton transfer plays an important role in this process \citep{Mahalingam_2008}. Afterward, the protein bleaches and returns to the dark state, completing what is called the photocycle. As the protein conformation changes to META, TM6 moves characteristically outward, activating the G-protein. In photoreceptor cells, the G-protein, transducin, activates a phosphodiesterase, which hydrolyzes cGMP into GMP, decreasing the activity of cGMP-gated cation channels and hyperpolarizing the cell \citep{Chabre_1989}.

In this study, I select 9 opsins expressed in the human nervous system: OPN1MW (UniProt ID: \href{https://www.uniprot.org/uniprotkb/P04001/entry}{P04001}), OPN1LW (UniProt ID: \href{https://www.uniprot.org/uniprotkb/P04000/entry}{P04001}), OPN1SW (UniProt ID: \href{https://www.uniprot.org/uniprotkb/P03999/entry}{P03999}), OPN2 (UniProt ID: \href{https://www.uniprot.org/uniprotkb/P08100/entry}{P04001}), OPN3 (UniProt ID: \href{https://www.uniprot.org/uniprotkb/Q9H1Y3/entry}{Q9H1Y3}), OPN4 (UniProt ID: \href{https://www.uniprot.org/uniprotkb/Q9UHM6/entry}{Q9UHM6}), OPN5 (UniProt ID: \href{https://www.uniprot.org/uniprotkb/Q6U736/entry}{Q6U736}), RGR (UniProt ID: \href{https://www.uniprot.org/uniprotkb/P47804/entry}{P47804}) and RRH (UniProt ID: \href{https://www.uniprot.org/uniprotkb/O14718/entry}{O14718}). They belong to several evolutionarily distinct families of animal opsins \citep{Terakita_2005}.

@@

\textbf{OPN2} (Opsin 2), also known as rhodopsin, is expressed in retinal rod photoreceptors and is responsible for vision at low light intensity \citep{Hubbard_1958}. Certain variants lead to autosomal recessive or autosomal dominant retinitis pigmentosa and congenital stationary night blindness \citep{Fanelli_2021}. OPN2 is a representative animal opsin, one of the first studied. In fact, bovine OPN2 is the first opsin to be sequenced \citep{Nathans_1984}, as well as the first GPCR whose crystal structure was resolved experimentally \citep{Palczeski_2000}. Many studies on the functional mechanisms of animal opsins also focus on OPN2. Consequently, I choose OPN2 to conduct a functional analysis in this study in order to further explore the effectiveness of the QTY code in redesigning retinylidene proteins.

@@

In this study, I select 3 microbial opsins: BACR (UniProt ID: \href{https://www.uniprot.org/uniprotkb/P02945/entry}{P02945}), BACH (UniProt ID: \href{https://www.uniprot.org/uniprotkb/B0R2U4/entry}{B0R2U4}), ChR2 (UniProt ID: \href{https://www.uniprot.org/uniprotkb/Q8RUT8/entry}{Q8RUT8}).

@@

Instead of taking a computational approach, I apply the QTY (Glutamine, Threonine, Tyrosine) code to systematically engineer water-soluble analogs with reduced hydrophobicity in membrane proteins. There are structural similarities between hydrophobic and polar amino acids: leucine (L) and glutamine (Q); isoleucine (I) / valine (V) and threonine (T); and phenylalanine (F) and tyrosine (Y), as can be observed on high-resolution electron density maps \citep{Zhang_2018, Zhang_2022, Tegler_2020}. This justifies the replacement of hydrophobic amino acids with polar ones. Zhang et al. initially applied the QTY code to design several detergent-free chemokines receptors, all of which retained structural thermal stability and native ligand-binding and enzymatic activities despite substantial changes to the transmembrane domain \citep{Zhang_2018, Tegler_2020}. Zhang's team also applied the QTY code to design water-soluble GPCRs, including chemokine receptors \citep{Zhang_2018, Qing_2019, Tegler_2020, Skuhersky_2021}, cytokine receptors \citep{Hao_2020}, and olfactory receptors \citep{Skuhersky_2021, Johnsson_2025}. In addition to GPCR, they used the QTY code to design water-soluble glucose transporters \citep{Smorodina_2022}, ABC transporters \citep{Pan_2024}, monoamine neurotransmitter transporters \citep{Karagol_Karagol_Alper_2024}, glutamate transporters \citep{Karagol_Karagol_Smorodina_2024, Karagol_2024}, mitochondrial megacomplex \citep{Chen_2025}, antibodies \citep{Li_Wang_2024}, potassium channels \citep{Smorodina_2024}, receptor kinases \citep{Li_Tang_2024}, transmembrane enzymes \citep{Chen_Pan_2025}, as well as bacterial enzymes with beta barrel structures \citep{Sajeev-Sheeja_2023, Sajeev-Sheeja_2024}. The QTY-designed water-soluble CXCR4 chemokine receptor was successfully used in biomimetic sensors \citep{Qing_2023}.

AlphaFold2 was released by Google DeepMind in July 2021 \citep{Jumper_2021}. It has greatly facilitated the study of QTY-designed protein variants. Subsequently, AlphaFold3 was released in May 2024, featuring improved architecture and improved efficiency \citep{Abramson_2024}. Furthermore, AlphaFold3 enabled the accurate prediction of complexes of multiple proteins, as well as complexes with modified residues, nucleic acids, ions, and certain ligands. QTY studies have made use of these new features \citep{Chen_2025, Johnsson_2025}.

@@

I predicted the structures of native proteins and their QTY variants using the AlphaFold3 website (\url{https://alphafoldserver.com}) \citep{Abramson_2024}. For microbial opsins, the dimers and trimers have identical subunits, i.e., are homodimers and homotrimers. In AlphaFold3, this was achieved by altering the number of protein copies.

@@

PDB files for native protein structures determined experimentally by X-ray diffraction or electron microscopy were taken from RCSB PDB: OPN2 (PDB ID: \href{https://www.rcsb.org/structure/5W0P}{5W0P}), BACR (PDB ID: \href{https://www.rcsb.org/structure/7XJC}{7XJC}), BACH (PDB ID: \href{https://www.rcsb.org/structure/2JAF}{2JAF}), ChR2 (PDB ID: \href{https://www.rcsb.org/structure/8ZAN}{8ZAN}). The AlphaFold3-predicted native and QTY variants were taken directly from the most probable predicted structure. Superpositions were performed via the command ``super'' in PyMOL (\url{https://pymol.org}). I removed unstructured loops at the N and C terminals for the sake of clarity.

@@

Regarding native OPN2, the protein-membrane system was constructed using the web-based membrane builder CHARMM-GUI and downloaded in GROMACS format \citep{Jo_2008, Wu_2014, Lee_2016}. The protein was centered in a rectangular box. The generated membrane models consisted of 40\% POPC, 40\% POPE, 10\% POPS, and 10\% cholesterol, which simulated a rod photoreceptor disk membrane where native OPN2 is usually expressed \citep{Albert_2005}. The system was solvated in TIP3P water with 150 mM NaCl. NaCl was used instead of KCl to simulate the environment of rod photoreceptor outer segments \citep{Govardovskii_1971}. 11-cis-retinal was manually added to the protein files. LINCS constraints were used for constraints, and Verlet integrator was used. Electrostatics was handled with Particle Mesh Ewald (PME), with both Coulomb and van der Waals interaction cutoffs set at 1.2 nm. The energy of the system was minimized using the steepest descent until the maximum forces converged below 1000 kJ/mol/nm. The standard six-step CHARMM-GUI NP$\gamma$T equilibration protocol \citep{Jo_2008} was used, with two 125-ps NVT equilibration simulations, and one 125-ps and three 500-ps NP$\gamma$T equilibration simulations. Temperature and pressure were maintained at 303.15 K and 1.0 bar, respectively, using the V-rescale thermostat and C-rescale barostat with surface-tension coupling. Following equilibration, a 10-ns production MD simulation was run. The retinal molecule was then manually switched to the all-trans state by rotating C13 to C15 and associated hydrogen and methyl groups by 180 degrees around the C11-C12 bond. This mimicked the effect of the absorption of a photon by 11-cis-retinal. Energy minimization and equilibration were repeated exactly as before, and then a 120-ns production MD simulation was run.

Regarding QTY-designed OPN2, I found that the helix 9 of the QTY analog flung up and adhered to the surface that was originally the transmembrane domain, which disrupted the normal conformation. Therefore, this helix was truncated. The system was constructed directly using GROMACS. The protein was centered in a rectangular box with dimensions equal to those of native OPN2. The equilibration involved a sequence of simulations: one 250-ps NVT equilibration, followed by one 125-ps and one 1500-ps NPT equilibrations, totaling an equilibration duration that matches that of the natural OPN2. All other configurations and parameters were identical to those of the original OPN2.

@@

The transmembrane segments of membrane proteins are hydrophobic in nature. This leads to a challenge to study their structure and function. Traditionally, detergent must be used to purify them. Zhang et al. considered a different approach, i.e., via systematic soluble protein design. There exist structural similarities, as can be observed on high-resolution electron density maps, between certain hydrophobic and hydrophilic amino acids: leucine (L) and glutamine (Q), isoleucine (I) or valine (V) and threonine (T), and phenylalanine (F) with tyrosine (Y). This fact enables systematic replacement of L with Q, I/V with T, and F with Y in all transmembrane segments of the membrane protein, which Zhang et al. named the QTY code. QTY analogs are less hydrophobic than native membrane proteins. Although their amino acid sequences are significantly changed, they still exhibit relatively preserved structure, isoelectric points (pI), and molecular weights (MW) compared to the native membrane proteins (Table~\ref{tb:characteristics}).

@@

The protein sequences of 9 human opsins and 3 microbial opsins were aligned with their QTY analogs (Figure~\ref{fig:sequences}). The QTY substitution resulted in an overall variation of the sequence of 15.48\% to 30.04\% and a variation of transmembrane domain of 35.53\% to 50.24\%. Despite changes in amino acid composition and sequence, the pI only experienced a slight change, between 0.00 and 0.16. This is because the amino acids Q, T, and Y have neutral charges and do not introduce additional charges to the system.  The MW increased slightly by a value between 0.04kDa and 0.60kDa. This is because the hydrophilic amino acids usually have nitrogen or oxygen atoms in place of carbon atoms in hydrophobic acids and thus may have higher mass: L (131.17Da) \textit{vs} Q (146.14Da); I (131.17Da) and V (117.15Da) \textit{vs} T (119.12Da); F (165.19Da) \textit{vs} Y (181.19Da).

@@

Although the chromophore, retinal, is bound to retinylidene proteins in an internal pocket, it is only incorporated into the protein after protein folding is complete. In other words, protein folding does not depend on the chromophore. Thus, it is justified to predict the ligand-free forms of the proteins. I used AlphaFold3 to predict the structure of the 9 native human opsins and their water-soluble QTY analogs. The structures superposed very well (Figure~\ref{fig:humansup} \textbf{a)} to \textbf{i)}). For OPN2, which has an available X-ray diffraction structure, I also superposed the experimentally determined structure with the AlphaFold3-predicted structures. Overall, the root mean square distances (RMSD) were small, from 0.307{\AA} to 0.611\AA, with only one exception of OPN2$^{\textrm{QTY}}$ \textit{vs} OPN2$^{\textrm{EXP}}$ (RMSD = 0.999\AA). This shows that the water-soluble QTY analogs are quite similar to native proteins in terms of structure.

All 9 of these human opsins belong to the same large family, i.e. animal opsins. However, they belong to different subfamilies and have structural differences from each other. In order to investigate the extent to which the water-soluble QTY analogs preserve these differences, I also conducted pairwise superpositions within all 9 native human opsins and within all 9 QTY analogs of human opsins (Figure~\ref{fig:humansup} \textbf{j)} and \textbf{k)}). As observed in the tables, there was a large degree of similarity between the RMSD between a pair of native proteins and that between the corresponding pair of QTY analogs. Therefore, I conclude that the structural differences are relatively well-preserved before and after the QTY substitution. However, whether the functional differences are preserved remains a problem to study.

@@

I used AlphaFold3 to predict the structure of monomers of the 3 native microbial opsins, their water-soluble QTY analogs. Since microbial opsins often have to form oligomers to be functional, I also predicted the structure of native and QTY variants of BACR trimers, BACH trimers, and ChR2 dimers. These predicted structures were also superposed with experimentally determined structures (X-ray diffraction or electron microscopy). The structures superposed very well (Figure~\ref{fig:microbialsup}). Overall, the root mean square differences (RMSD) were small, with the highest value being 0.685\AA. This shows that the water-soluble QTY analogs are likely capable of forming oligomers with structures similar to those of native proteins.

@@

The hydrophobic portions of the protein surface are represented in yellow (Figure~\ref{fig:hydrophobicity}). In the native proteins, there are large hydrophobic patches in the transmembrane domains, which are embedded within the lipid bilayer. The nonpolar and hydrophobic amino acids, including leucine (L), isoleucine (I), valine (V), phenylalanine (F), alanine (A), methionine (M), and tryptophan (W), interact with the hydrophobic hydrocarbon chains and expel water.

I investigated the surface hydrophobicity of water-soluble QTY-analogs in comparison with the native proteins. After applying the QTY code to respectively replace the hydrophobic amino acids L, I/V, and F with hydrophilic amino acids glutamine (Q), threonine (T), and tyrosine (Y), the hydrophobic surface areas were significantly reduced. More importantly, since the electron density maps of the corresponding amino acids are similar, the alpha helices in QTY analogs had a similar contour to those in the native proteins. The QTY analogs successfully retained its structural integrity and stability.

@@

AlphaFold3 is easily accessible online (\url{https://alphafoldserver.com}), allowing users to make 30 predictions a day at present. The structures of the QTY analogs were predicted using the AlphaFold3 server, which was free of charge, and the results were produced within a few minutes.

DeepMind also collaborated with the EBI to make over 214 million predicted protein structures available through the AlphaFold Protein Structure Database (\url{https://alphafold.ebi.ac.uk}). This number is continuously expanding, with the quality of predictions further improving with the advent of AlphaFold 3.

However, AlphaFold3 still has limitations that have been encountered in this study. Although AlphaFold3 allows the structural prediction of complexes that include small-molecule ligands, it only supports a small range of ligands, which does not include retinal. I was therefore unable to predict the structure of chromophore-bound states of retinylidene proteins with AlphaFold3. Consequently, I chose another approach, namely molecular dynamics simulation, in order to investigate the behavior of QTY opsin analogs when they are bound to the chromophore.

@@

QTY substitutions do not introduce changes to the essential NPxxY motif or the retinal-binding lysine of human retinylidene proteins. Therefore, I proposed that QTY analogs of human opsins may have conserved functions and carried out a molecular dynamics (MD) simulation. I recorded and analyzed the behavior of native OPN2 and its QTY analog before and after the 11-cis to all-trans isomerization. Please note that my aim is not to simulate the whole photocycle of OPN2, which takes longer than 1ms \citep{Fanelli_2021}, but rather to find proof that the QTY analog conserves the function of being activated by isomerization of retinal.

The MD simulation provided evidence of conformational change in both native and QTY analogs (Figure~\ref{fig:conformation}). The native protein experienced a slower change from its original conformation, while the QTY analog showed more fluctuations and sudden changes. Under further inspection, this phenomenon might be due to the entrance of water molecules into the QTY analog. In addition, retinal molecule had a greater degree of change in conformation in the native protein than in the QTY analog. Nevertheless, the retinal in the QTY analog was still able to cause changes to surrounding residues, sometimes doing so via intervening water molecules. It may be noted that the RMSD was still fluctuating, which means that the protein had not yet stabailized. This was expected, since the OPN2 was still in the process of transition from BATH to LUMI state, and this provided evidence for the activation of the protein in response to isomerization of retinal.

Zooming in on the retinal-binding pocket, I observed that retinal is held within its binding pocket by several forces in both the native and QTY analog: the counterions of the retinal Schiff base (RSB) (113E and 181E), the hydrophobic interactions with neighboring residues (for example, 265W, 268Y, and the 187--189 beta strand), and, sometimes, steric collisions (Figure~\ref{fig:binding}). Since retinal is a ligand that binds in an internal pocket, I was unable to calculate the Gibbs free energy of its binding. Nevertheless, the distributions of interaction energies were found to be similar in native OPN2 and the water-soluble QTY analog, except for small differences: native OPN2 experienced a change in interaction energy at about $t=50$ns due to the entrance of water near 113E; the QTY analog had more negative interaction energy at residues 208 and 212 due to phenylalanine (F) to tyrosine (Y) substitutions. Apart from these, the similarity was a strong suggestion of functional conservation. Finally, I observed very different hydrogen bonding and water molecule patterns inside the binding pocket, which was mainly due to several F to Y substitutions in that region. This might cause retinal to interact with adjacent residues in a subtly different way, and might lead to a different absorption peak since the environment of the RSB, i.e., the binding pocket, determines the spectral characteristic of opsins \citep{Fenno_2011}.
